Demanding resources: Converging perceptions of challenge and resource characteristics

### 2023-07-10 SIOP2024/SIOP2024convergence.Rmd
SIOP Edits

```
The Occupational Information Network (O\*NET; onetonline.org) contains a reasonably comprehensive description of occupations [@peterson2001understanding] which subject matter experts inform via surveys. We utilize a set of O\*Net job context and activity statements contained in these surveys to explore a series of research questions about how people perceive resources, challenges, and hindrances within their jobs. There are a number of advantage to using this standard set of statements, but the primary one here lies in our ability to consider a wide range of job characteristics, with a focus on individual perceptions of each rather than limiting ourselves to a more generic set of demands or those of a limited scope.

@@

At the most broad level, we can question whether employees actually distinguish between challenges and hindrances as the theory describes. Research suggests that people can and do perceive them differently. For example, @bakker2013weekly found that perceived work pressure can be classified as a hindrance demand, while the requirement to express emotions is a challenge demand. @webster2011extending also considered three common work characteristics including workload, role ambiguity, and role conflict. Interestingly, they found that while each could be identified *primarily* as a challenge or hindrance, employees could also appraise that a characteristic is simultaneously both a challenge and hindrance. However, this list of demands (workload, role ambiguity, and role conflict) is somewhat narrow. Using a very wide range of O\*Net job characteristics regarding job context and activities, we explore this question:

*RQ1 is a good idea for A DIFFERENT PAPER. Do ICC(1,1)*
*RQ1*: Do people distinguish between challenge and hindrance stressors?

*Paired samples t-tests for challenge/hindrance by item?* - I like this first idea. could present it by the 7 different categories.
*Could look*
*Or percent agreement?* 
*Maybe do item-level correlations between challenge/hindrance rating and present percent positively related, and negatively related (negative would be supportive, I think)*

The next question focuses on whether challenges (positive stressors per the definition above) might actually be positively related to resources. Per the job demands-resources theory [@bakker2017job] resources are defined as physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. It stands to reason that a challenging demand may be considered a resource.

*RQ1*: Are job characteristics perceived as challenges are also commonly viewed as resources?  
Although we argue that the same job characteristic may be perceived as both a challenge and a resource, it is also likely that some characteristics are less likely to be viewed as mutually complementary as others. For example, a physically strenuous job requirement such as "carrying heavy objects" would be less likely to be viewed both as a challenge and a resource whereas a structural characteristic such as "negotiating work schedules" may very well be viewed (likely in different circumstances) to be both a control-oriented resource as well as a challenge. O\*Net has different levels of abstraction with regard to the nature of job characteristics. We will be exploring a mid-level abstraction with seven different characteristic scales.

@@

We evaluate agreement across perceptions of job characteristics regarding their characterization as resource, challenge, and hindrance (Bakker & Demerouti, 2017; Bakker et al., 2003; Demerouti et al., 2001). To capture an effectively exhaustive list of characteristics that apply to, theoretically, *every* possible job, we consult the  unifying framework of O\*Net.

@@

Our survey was informed by 98 statements taken from job characteristic descriptors located within O\*Net’s classification of “work activities”: 1) Information Input (5 statements), 2) Interacting with Others (17 statements), 3) Mental Processes (10 statements), and 4) Work Output (9 statements) and “work context” groupings: 1) Interpersonal Relationships (14 statements), 2) Physical Work Conditions (30 statements), and 3) Structural Job Characteristics (13 statements).

The O\*Net descriptors are written in a similar manner to a task statement presented within a job analysis, but the level of abstraction is closer to "responsibility" than task. For example, the descriptor for "level of competition", which is an element of the "structural job characteristics" grouping, is *...to what extent does this job require the worker to compete or to be aware of competitive pressures?* Other than minor grammatical editing (for example, changing “the worker” to “you”), we retained the O\*Net wording for our item stems. We also retained O\*Net’s response scales, several of which were semantically unique across items, but all shared the same 5-point scale. It would likely NOT be considered controversial to referred to these as "effectively" Likert-type response scales.

@@

Following IRB approval, respondents were recruited through Prolific. After providing consent, respondents provided ratings of whether or not each of the 98 O\*Net characteristics were relevant for the respondent's work. Subsequently, each respondent who agreed that an element had at least some relevance to their job was then also asked to rate that element in terms of perception of resource, challenge, or hindrance. The total number of items on the survey was less than 392 (98 characteristics x 4 administrations - the first for relevance and the remaining 3 for resources, challenges, and hindrances) because we did not ask for demand and resource evaluations for 14 O\*Net characteristics that we projected would have very low frequency of endorsement across respondents (one excluded characteristic, for example, was *...the extent to which the worker is exposed to radiation on the job*). Furthermore, not every respondent indicated that the same characteristics were relevant to them, therefore each respondent had a unique set of responses. Participants were compensated for their participation in this study in the amount of six dollars through Prolific.

@@

All analyses are focused on characteristics of work that were rated as being "relevant" to the respondents' job. Upon confirming that a work characteristic was relevant, respondents then also rated the extent to which that characteristic was percieved as a resources, challenge, and hindrance.

@@

Hypothesis 1 predicted a positive association between total resources and total challenge demands. Table \@ref(tab:cortab) shows a very high association between resources and challenges (*r* = .86). These associations, however, are only capturing the relationships between these demands and resources in *sheer volume*. That is, table \@ref(tab:cortab) operationalized each variable as the sheer number of resources, hindrances, or challenges that a respondent indicated were present within their job. This correlational analysis simply implies that workers who experience more resources also perceive greater challenges. The associations between resources and hindrances (*r* = .23) and challenges and hindrances (*r* = .22) were also significantly positive with moderate magnitude associations, suggesting that some of this "sheer volume" may be capturing job complexity (that is, the more complex the job, the more characteristics are relevant, and therefore the more likely it is to have *more* challenges as well as *more* hindrances as well as *more* resources). Although we did not address job complexity as a moderator in the current paper, we do plan to do so in future investigations. Also, take note of the average numbers of resources, challenges, and hindrances cited by our sample, where these respondents generally experienced fewer hindrances in their jobs (*M* = 13.09) than both resources and challenges.

@@

## lotsa variance in resource-challenge agreement. Next look for either content (O*Net category) or contextual (job) moderator

```
## Convergence (same characteristic).

@@

Figure \@ref(fig:recchall) explores the possibility of moderation by *type of characteristic rated* for the resource-challenge convergence. Here we categorized each characteristic by its O\*Net "scale" (one of seven), and the graph shows greater consistency across certain characteristics (for example, *Mental Processes* or *Interacting with Others*) and less convergence across other *types* of job activities (for example, *Physical* characteristics). A repeated-measures ANOVA retaining these scales as 7 different levels of a within-subjects' independent variables yielded a treatment effect of $F_{(6, 3,402)}$ = 613.5, *p* < .001 (the subjects' effect was $F_{(567, 3402)}$ = 6.13, *p* < .001.

```{r workcat, fig.cap="Percent agreement (resource and challenge) across O*Net occupational categories (previous was across 84 O/*Net items - this is across 568 individual workers).", fig.height=10}

@@

forworkers2 <- anova[,c(2:85)]
dataworker <- cbind(forworkers2,workers$O.Net.Job.Family.Categories)    ## to repeat O*Net analyses
                                     ## This gets worker info into convergence file

@@

As can be seen in Figure \@ref(fig:workcat), the moderation seen across O\*Net dimension was not paralleled across *job classification*. Here, there was wide variability in convergence, regardless of *type* of job. The omnibus ANOVA did breach statistical significance ($F_{(7, 560)}$ = `r summary(workeraov)[[1]][1,4]`, *p* = `r summary(workeraov)[[1]][1,5]`), however, the practical effect was very small ($\eta^2$ = `r effect.size[1,1]`). The only contrast approaching statistical significance via Tukey's HSD specification (*p* = .09) was the contrast between Physical Labor (*M* = 0.49) and Business and Financial Operations (*M* = 0.58). Additionally, a Levene's test for homogeneity of variance was non-significant (Levene's = `r levenes[1,2]`, *p* = `r levenes[1,3]`). [should be .6539 and .7112]

>>>>

```
The Occupational Information Network (O\*NET; onetonline.org) contains a reasonably comprehensive description of occupations [@peterson2001understanding] which subject matter experts inform via surveys. We utilize a set of O\*NET job context and activity statements contained in these surveys to explore a series of research questions about how people perceive resources, challenges, and hindrances within their jobs. There are a number of advantage to using this standard set of statements, but the primary one here lies in our ability to consider a wide range of job characteristics, with a focus on individual perceptions of each rather than limiting ourselves to a more generic set of demands or those of a limited scope.

@@

At the most broad level, we can question whether employees actually distinguish between challenges and hindrances as the theory describes. Research suggests that people can and do perceive them differently. For example, @bakker2013weekly found that perceived work pressure can be classified as a hindrance demand, while the requirement to express emotions is a challenge demand. @webster2011extending also considered three common work characteristics including workload, role ambiguity, and role conflict. Interestingly, they found that while each could be identified *primarily* as a challenge or hindrance, employees could also appraise that a characteristic is simultaneously both a challenge and hindrance. However, this list of demands (workload, role ambiguity, and role conflict) is somewhat narrow. Using a very wide range of O\*NET job characteristics regarding job context and activities, we explore whether challenges (positive stressors per the definition above) might actually be positively related to resources. Per the job demands-resources theory [@bakker2017job] resources are defined as physical, psychological, social, or organizational aspects of the job that may help an employee achieve work goals, reduce job demands, or promote personal growth and development [@demerouti2001job]. It stands to reason that a challenging demand may be considered a resource.

*RQ1*: Do jobs with many resources also perceived to have many challenges?

Although we argue that the same job characteristic may be perceived as both a challenge and a resource, it is also likely that some characteristics are less likely to be viewed as mutually complementary as others. For example, a physically strenuous job requirement such as "carrying heavy objects" would be less likely to be viewed both as a challenge and a resource whereas a structural characteristic such as "negotiating work schedules" may very well be viewed (likely in different circumstances) to be both a control-oriented resource as well as a challenge. O\*NET has different levels of abstraction with regard to the nature of job characteristics. We will be exploring a mid-level abstraction with seven different characteristic scales.

@@

We evaluate agreement across perceptions of job characteristics regarding their characterization as resource, challenge, and hindrance (Bakker & Demerouti, 2017; Bakker et al., 2003; Demerouti et al., 2001). To capture an effectively exhaustive list of characteristics that apply to, theoretically, *every* possible job, we consult the  unifying framework of O\*NET.

@@

Our survey was informed by 98 statements taken from job characteristic descriptors located within O\*NET’s classification of “work activities”: 1) Information Input (5 statements), 2) Interacting with Others (17 statements), 3) Mental Processes (10 statements), and 4) Work Output (9 statements) and “work context” groupings: 1) Interpersonal Relationships (14 statements), 2) Physical Work Conditions (30 statements), and 3) Structural Job Characteristics (13 statements).

The O\*NET descriptors are written in a similar manner to a task statement presented within a job analysis, but the level of abstraction is closer to "responsibility" than task. For example, the descriptor for "level of competition", which is an element of the "structural job characteristics" grouping, is *How often do you have to perform public speaking in this job??* Other than minor grammatical editing (for example, changing “the worker” to “you” where applicable), we retained the O\*NET wording for our item stems. We also retained O\*NET’s response scales, several of which were semantically unique across items, but all shared the same 5-point scale. It would likely NOT be considered controversial to referred to these as "effectively" Likert-type response scales.

@@

Respondents were recruited through Prolific. After providing consent, respondents provided ratings of whether or not each of the 98 O\*NET characteristics were relevant for the respondent's primary work role. Subsequently, each respondent who agreed that an element had at least some relevance to their job was then also asked to rate that element in terms of perception of resource, challenge, and hindrance. The total number of items on the survey was less than 392 (98 characteristics x 4 administrations - the first for relevance and the remaining 3 for resources, challenges, and hindrances) because we did not ask for demand and resource evaluations for 14 O\*NET characteristics that we projected would have very low frequency of endorsement across respondents (one excluded characteristic, for example, was *...the extent to which the worker is exposed to radiation on the job*). Furthermore, not every respondent indicated that the same characteristics were relevant to them. Therefore, each respondent had a unique set of administered items. Participants were compensated for their participation in this study in the amount of six dollars through Prolific.

@@

All analyses are focused on characteristics of work that were rated as being "relevant" to the respondents' primary job. Upon confirming that a work characteristic was relevant, respondents then also rated the extent to which that characteristic was perceived as a resources, challenge, and hindrance.

@@

Research Question 1 explored whether jobs with many challenges have many resources. The total number of undifferentiated challenges and resources are positively related. Table \@ref(tab:cortab) shows a very high association between resources and challenges (*r* = .86). These associations, however, are only capturing the relationships between these demands and resources in *sheer volume*. That is, Table \@ref(tab:cortab) operationalized each variable as the sheer number of resources, hindrances, or challenges that a respondent indicated were present within their job. This correlational analysis simply implies that workers who experience more resources also perceive greater challenges. The associations between resources and hindrances (*r* = .23) and challenges and hindrances (*r* = .22) were also significantly positive with moderate magnitude associations, suggesting that some of this "sheer volume" may be capturing job complexity (that is, the more complex the job, the more characteristics are relevant, and therefore the more likely it is to have *more* challenges as well as *more* hindrances as well as *more* resources). Although we did not address job complexity as a moderator in the current paper, we do plan to do so in future investigations. Also, take note of the average numbers of resources, challenges, and hindrances cited by our sample, where these respondents generally experienced fewer hindrances in their jobs (*M* = 13.09) than both resources and challenges.

## Convergence (same characteristic)

@@

## lotsa variance in resource-challenge agreement. Next look for either content (O*NET category) or contextual (job) moderator

@@

Figure \@ref(fig:recchall) explores the possibility of moderation by *type of characteristic rated* for the resource-challenge convergence. Here we categorized each characteristic by its O\*NET "scale" (one of seven), and the graph shows greater consistency across certain characteristics (for example, *Mental Processes* or *Interacting with Others*) and less convergence across other *types* of job activities (for example, *Physical* characteristics). A repeated-measures ANOVA retaining these scales as 7 different levels of a within-subjects' independent variables yielded a treatment effect of $F_{(6, 3,402)}$ = 613.5, *p* < .001 (the subjects' effect was $F_{(567, 3402)}$ = 6.13, *p* < .001.

```{r workcat, fig.cap="Percent agreement (resource and challenge) across O*NET occupational categories (previous was across 84 O/*Net items - this is across 568 individual workers).", fig.height=10}

@@

forworkers2 <- anova[,c(2:85)]
dataworker <- cbind(forworkers2,workers$O.Net.Job.Family.Categories)    ## to repeat O*NET analyses
                                     ## This gets worker info into convergence file

@@

As can be seen in Figure \@ref(fig:workcat), the moderation seen across O\*NET dimension was not paralleled across *job classification*. Here, there was wide variability in convergence, regardless of *type* of job. The omnibus ANOVA did breach statistical significance ($F_{(7, 560)}$ = `r summary(workeraov)[[1]][1,4]`, *p* = `r summary(workeraov)[[1]][1,5]`), however, the practical effect was very small ($\eta^2$ = `r effect.size[1,1]`). The only contrast approaching statistical significance via Tukey's HSD specification (*p* = .09) was the contrast between Physical Labor (*M* = 0.49) and Business and Financial Operations (*M* = 0.58). Additionally, a Levene's test for homogeneity of variance was non-significant (Levene's = `r levenes[1,2]`, *p* = `r levenes[1,3]`). [should be .6539 and .7112]
